A cautionary note on the use of Ornstein Uhlenbeck and other models in macroevolutionary studies

### 2015-04-13 Manuscript/ohyou.tex
Edits to 3 uses, thinking to change to two uses...

  The popularity of the OU model has grown extensively in recent years (Fig. \ref{figure.literature}); over 2500 ecology, evolution and palaeontology papers containing the phrase ``Ornstein Uhlenbeck'' were published between 2012 and 2014 (Google Scholar search 15th March 2015; see Supplementary Material). % Need to add ref to supp mat
  This may partly be because these models are now easy to implement apply via packages in R \citep[e.g. ouch, GEIGER and OUwie;][]{Butler:2004aa,Harmon:2008aa,beaulieu2012ouwie}. 
  Additionally, although the OU model is pattern-based, it has a number of attractive biological interpretations. 
  For example, fit to an OU model is used as evidence for the action of processes such as phylogenetic niche conservatism, convergent evolution and stabilizing selection \citep[e.g.,][]{Wiens:2010aa,christin2013anatomical,ingram2013surface}. 
  Unfortunately, although widely used, the properties of the OU model are poorly understood leading to the potential for inappropriate use and misinterpretation of results.

@@

  Uses of the OU model can be divided into three broad categories. % improve citations in this section?
    \begin{enumerate}
      \item \textbf{Modeling the evolution of a continuous character.}\\ 
            Generally in these analyses, multiple models of evolution are fit to the same continuous character \citep[e.g.,][]{Harmon:2010aa}. Model selection is then used to determine which model best fits the data. 
            If the OU model is favoured, this is often used as evidence of evolutionary constraints, stabilising selection, or phylogenetic niche conservatism \citep[e.g.,][]{Harmon:2010aa,cooper2010body,Wiens:2010aa,christin2013anatomical}.
            OU models with different numbers of optima may also be included in the set of models considered (see below).
      \item \textbf{Modeling the effects of different hypothesised adaptive regimes on a continuous character.}\\
            OU models can be fit with one optimal trait value, or multiple different optima \citep{Butler:2004aa}. 
            The latter represents evolution under multiple selective regimes, and is probably more biologically realistic. 
            In these analyses, OU models with different numbers of optima are fitted to the same continuous character, then model selection is used to determine what number of selection regimes best fits the evolution of the trait \citep[e.g.,][]{beaulieu2012modeling}.
      \item \textbf{Controlling for phylogeny in correlative analyses.}\\ 
            Phylogenetic generalised least squares (PGLS) models incorporate information about the relationships among species into the error term of a generalised least squares model. 
            This error term generally consists of a variance-covariance matrix of the phylogeny, but various transformations are used (e.g., Pagel's $\lambda$; \citealp{Pagel:1997aa}) to improve the fit of the model to the data. 
            In these analyses, the OU model is used to transform the tree and the variance-covariance matrix.
            This is rarely interpreted as corresponding to any kind of process, it just improves the fit of the PGLS models \citep[e.g.,][]{blankers2012ecological}.
    \end{enumerate}

@@

    \begin{figure}[h]
      \centering
      \includegraphics[width=10cm, height=10cm, keepaspectratio=true]{Figures/PapersThruTime.png}
      \caption{The number of ecology, evolutionary biology and palaeontology papers published between 2005 and 2014 containing the phrase ``Ornstein-Uhlenbeck'', as a proportion of the total number of ecology, evolutionary biology or palaeontology papers published that year. See Supplementary Material for details %linkto supp mat
      }
      \label{figure.uses}
    \end{figure}

@@

    When $\alpha$ is close to zero, evolution is approximately Brownian, then as $\alpha$ gets larger the non-Brownian behaviour of the model starts to become apparent. 
    Eventually, when $\alpha$ is really large, all imprint of history is lost and the trait evolution is essentially a rapid burst at the present (FIGURE 2). % Fix the figure or send to supplemental
    We discuss how to interpret values of $\alpha$ below.

>>>>

  The popularity of the OU model has grown extensively in recent years (Fig. \ref{figure.literature}); over 2500 ecology, evolution and palaeontology papers containing the phrase ``Ornstein Uhlenbeck'' were published between 2012 and 2014 (Google Scholar search 15th March 2015; see Supplementary Material). % Need to add ref to supp mat
  This may partly be because these models are now easy to implement apply via packages in R \citep[e.g. ouch, GEIGER and OUwie;][]{Butler:2004aa,Harmon:2008aa,beaulieu2012ouwie}. 
  Additionally, although the OU model is pattern-based, it has a number of attractive biological interpretations. 
  For example, fit to an OU model is used as evidence for the action of processes such as phylogenetic niche conservatism, convergent evolution and stabilising selection \citep[e.g.,][]{Wiens:2010aa,christin2013anatomical,ingram2013surface}. 
  Unfortunately, although widely used, the properties of the OU model are poorly understood leading to the potential for inappropriate use and misinterpretation of results.

@@

  Uses of the OU model can be divided into three broad categories. % improve citations in this section? % fix up numbers. Argggggggghhhhhhhhhhh.
    \begin{enumerate}
      \item \textbf{Modeling the evolution of a continuous character (N = 58).}\\ 
            Generally in these analyses, multiple models of evolution are fit to the same continuous character \citep[e.g.,][]{Harmon:2010aa}. Model selection is then used to determine which model best fits the data. 
            If the OU model is favoured, this is often used as evidence of evolutionary constraints, stabilising selection, or phylogenetic niche conservatism \citep[e.g.,][]{Harmon:2010aa,cooper2010body,Wiens:2010aa,christin2013anatomical}.
      \item \textbf{Modeling the effects of different hypothesised adaptive regimes on a continuous character (N = 27).}\\
            OU models can be fit with one optimal trait value, or multiple different optima \citep{Butler:2004aa}. 
            The latter represents evolution under multiple selective regimes, and is probably more biologically realistic. 
            In these analyses, OU models with different numbers of optima are fitted to the same continuous character, then model selection is used to determine what number of selection regimes best fits the evolution of the trait \citep[e.g.,][]{beaulieu2012modeling}.
            Note that the first and second uses of the OU model are often combined, so that OU models with various numbers of optima are included in the pool of models being tested (N = 27).
      \item \textbf{Controlling for phylogeny in correlative analyses (N = 35).}\\ 
            Phylogenetic generalised least squares (PGLS) models incorporate information about the relationships among species into the error term of a generalised least squares model. 
            This error term generally consists of a variance-covariance matrix of the phylogeny, but various transformations are used (e.g., Pagel's $\lambda$; \citealp{Pagel:1997aa}) to improve the fit of the model to the data. 
            In these analyses, the OU model is used to transform the tree and the variance-covariance matrix.
            This is rarely interpreted as corresponding to any kind of process, it just improves the fit of the PGLS models \citep[e.g.,][]{blankers2012ecological}.
    \end{enumerate}

@@

    \begin{table}[h]
      \begin{centering}
        \caption{The most common uses of Ornstein Uhlenbeck models in ecology, evolutionary biology and palaeontology papers published between 2005 and 2013. See Supplementary Material for details linkto supp mat.}
        \bigskip  
          \begin{tabular}{p{8cm}lc}
            \hline
            Use of OU model & Optima & Number of papers\\
            \hline
            Ancestral state reconstruction & single & 8\\
            & multiple & 2\\
            Convergent evolution & single & 0\\
            & multiple & 2\\
            Model of evolution & single & 31\\
            & multiple & 27\\
            Phylogenetic generalised least squares & single & 35\\
            & multiple & 0\\
            Other & single & 5\\
            & multiple & 5\\
            Total & single &  36\\
            & multiple & 79\\
            \hline
            \label{table.uses}
          \end{tabular}
    \end{table}

@@

    When $\alpha$ is close to zero, evolution is approximately Brownian, then as $\alpha$ gets larger the non-Brownian behaviour of the model starts to become apparent. 
    Eventually, when $\alpha$ is really large, all imprint of history is lost and the trait evolution is essentially a rapid burst at the present (FIGURE 2). % Fix the figure or send to supplemental

    %$\alpha$ can most easily be interpreted by using it to estimate the ``phylogenetic half-life'' ($t_\frac{1}{2}$) of a trait, i.e., the time it takes for a species entering a new niche to evolve halfway toward its new expected optimum \citep{hansen1997stabilizing}, as follows:

      \begin{equation}
        t_\frac{1}{2} = \frac{ln(2)}{\alpha}
        \label{equation:halflife}
      \end{equation}

    %If $t_\frac{1}{2}$ is short relative to the branch lengths of the phylogeny, adaptation is quick, residual phylogenetic correlations are weak, and there is little influence of the past. This is sometimes referred to as ``phylogenetic inertia'' but we dislike this term
